---
title: Midterm Exam (2025)
author: Your Name
date: '2025-03-04'
categories:
- Exam
- Week07
type: exam
image: ../../figs/exam.svg
pub-date: '2023-03-05'
---


The United States Geological Survey (USGS) continuously monitors earth quakes and makes the information available to the public in a [data feed](https://earthquake.usgs.gov/earthquakes/feed/v1.0/csv.php). 
The dataset [all-month.csv](all_month.csv) included in this repository was pulled from USGS on Monday Mar 3, 2025. It contains all worldwide earth quakes  for the last 30 days.


An detailed description of all the variables in the `all_month` data (and more) is available from in the [ComCat Documentation](https://earthquake.usgs.gov/data/comcat/index.php)

For all of the following questions, provide answers and include your code. You can work in R or in python or in a mix of the two. 
Make sure that the final document you submit renders without errors. In the rendered document avoid any lengthy output. 


1. Load the data into the object `eq` using R or python. Make sure that you are working with an object with 9815   rows and 22 columns.

```{r}
# code chunk for R
```

```{python}
# python code chunk
```

2. Read the ComCat Documentation for the variables `time`, `latitude`, `longitude`, `depth`, `mag`, `magType`, `nst`, `id`, and `place`. Ensure that the format of the variables matches the description.  

If you work on R, use the `lubridate` package  to convert the variable `time` to a time object. If you work in python, use the `datetime` package.

```{r message = FALSE}
# R chunk
library(lubridate)
```

```{python}
# python chunk
import datetime

```

3. Include each of the variables `time`, `latitude`, `longitude`, `depth`, `mag`, `magType`, and `nst` in a data chart. Summarise the main finding(s) in each chart in 1-2 sentences. 

```{r}

```

```{python}

```

4. What is the time frame under consideration? How many earthquakes were there (show that the number of earthquakes is NOT equal to the number of rows in the dataset)?
When and where was the strongest earthquake? What was its magnitude?

```{r}

```

```{python}

```


5. `depth` is an estimate (in kilometers) of the depth in which an earthquake begins to rupture. The documentation suggests that in some cases default depths are used (when there is no further information available). Based on chart(s), argue which values are used for default depths and where? 

```{r}

```

```{python}

```

6. The variable `nst` contains the information on the number of seismic stations used to determine an earthquake's location.

  - What is the average number of seismic stations used to determine the location of an earthquake? 
  - What is the minimum number of stations? 
  - 1,592 earthquakes do not have a recorded number of stations. What else do these earthquakes have in common?


```{r}

```

```{python}

```

7. By introducing a suitable grouping variable, create a data set `daily_summary` that contains the daily summary values of the following statistics:

  - the total number of earthquakes 
  - the number of earthquakes with a magnitude of 4 or higher
  - the average magnitude of earthquakes
  - the location, place and magnitude of the strongest earthquake 

Based on the values in `daily_summary` discuss the statements: ***On average, there are about 50 earthquakes globally with a magnitude of at least 4***, ***days with more than 750 seismic events are not uncommon.***


8. "The Ring of fire" is a horseshoe-shaped belt along the rim of the Pacific (tectonic) Plate reaching from New Zealand to the tip of South America. The Ring of Fire contains hundreds of active and dormant volcanoes and is the location of a lot of seismic activity. 

    Introduce a transformation of the longitude variable `longitude_rof` to the data set that allows you to visualize the horseshoe shape of the ring of fire. (draw that chart)


```{r}

```

```{python}

```

9. Update the data set: write a function `update_eq` with arguments `data`, `url` and `verbose` that downloads the most recent 30 days of earthquake activity and incorporates that information into the provided data set `data`. 
Make sure to include each earthquake only once (in case there are multiple, choose the most recent update given by `updated`). If `verbose` is set to true, provide a summary of the changes made to the data set of the form (replace the stars by the corresponding numbers): 

    number of earthquakes added: *** 
    
    number of earthquakes updated: *** 

Call your function with the `eq` object set as the data, a `url` of https://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/all_month.csv
in the verbose mode. 
Save the resulting data set as `all_month_updated.csv`. Include the function, the call to your function and the updated file in your repository.


```{r}

```

```{python}

```
